cleaning script for foodborne_outbreaks.csv

### 2023-02-23 code/processing_code/processingfile_v1.qmd
4

# Setup
```{r}
library(readr) #for loading csv files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(dlookr)
```

@@

* Year, Month (date)
* Sate, Location (geographic information)
* Food, Ingredient
*Species, Serotype/Genotype
*Status

Following variables are outcomes:
*illnesses
*Hospitalizations
*Fatalities

@@

First, we will remove the variable "ingredient" and "serotype" since over 80% of data is missing. 
Does that mean it should be a numeric 60? It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby).
Since we don't know how to fix this, we might decide to remove the person. This "sixty" entry also turned all Height entries into characters instead of numeric. That conversion to character also means that our summary function isn't very meaningful. So let's fix that first.

```{r}
d1=
```

>>>>

# Setup
```{r}
library(tidyverse)
library(readr) #for loading csv files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(dlookr)
```

@@

* Year, Month (numeric)
* State, Location (geographic information)
* Food, -Ingredient
*Species, -Serotype/Genotype
*Status

Following variables are outcomes:
*illnesses
*Hospitalizations
*Fatalities
main outcome: illnesses

@@

First, we will remove the variable "ingredient" and "serotype" since over 80% of data is missing. 
Does that mean it should be a numeric 60? It somehow doesn't make sense since the weight is 60kg, which can't happen for a 60cm person (a baby).
Since we don't know how to fix this, we might decide to remove the person. This "sixty" entry also turned all Height entries into characters instead of numeric. That conversion to character also means that our summary function isn't very meaningful. So let's fix that first.
#potential code for EDA
#change month to numeric data
```{r}
rawdata1%>%filter(grepl("Confirmed", Status))
unique(rawdata1$Food)
d2=rawdata1%>%
  filter(!is.na(Food))%>%
  mutate(Food=fct_lump(Food,10))%>%
  count(Food)%>%
  arrange(desc(n))
d2
d3=rawdata1%>%
  filter(!is.na(Species))%>%
  mutate(Food=fct_collapse(Food,Norovirus=c("Norovirus","Norovirus genogroup I","Norovirus genogroup II")))%>%
  mutate(Food=fct_lump(Species,10))%>%
  count(Species)%>%
  arrange(desc(n))
d3
```
